---
title: "IML assignment 1"
author: "Emil Skindersø, Christopher Fjeldberg Jensen and Hans Vinther-Larsen"
date: "2024-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE,warning=FALSE}
library(tidyverse)
source("utils.R")
loadData()
```
# Introduction

can be found on github at https://github.com/HVinther/IML_assignment_1

# Data exploration, encoding and weighting

The dataset is a subset of the freMPL datasets from the CASdatasets library. 

## Encoding of dates and handling of missing values

Two <Date> columns are present in the dataset, RecordBeg and RecordEnd representing the beginning and end dates of recording. Several choices have to be made in the handling of these variables. First of, the `as_task` functions from `mlr3`cannot handle the <Date> datatype. A quick reencoding as `POSIXct`overcomes this initial hurdle - this will however need further encoding, as most models cannot handle the `POSIXct` class directly, but this can now be done inside a GraphLearner object. 

Before settling on an encoding strategy a second issue has to be considered. The RecordEnd variable has an abundance of missing values, in both the full dataset and in the training dataset about $46\%$ of the observations of this variable are missing. In particular these seem to be a result of right censoring, as all observations are within the year 2004, these particular observations are those, where the record has not ended within the year. Imputing these variables based on the nonmissing would expectedly result in an under estimation of their values.

We suggest and investigate the following encoding strategies for these variables:

A numerical encoding,imputing using lastest observed date and indicating censoring

```{r, eval = FALSE}
# Interpretes RecordEnd as being right censored.
# Thus imputes using the maximal observed time, 
# and adds column indicating censoring.
po_RecEnd_rc<-po(
  "mutate",
  mutation = list(
    RecordEnd = ~as.numeric(ifelse(is.na(RecordEnd),
                                   max(na.omit(RecordEnd)),
                                   RecordEnd)),
    RecordEnd_censored = ~is.na(RecordEnd)
  ),
  id = "RecEnd_rc"
)

# Reencodes RecordBeg (POSIXct) to numeric
po_RecBeg_num<-po(
  "mutate",
  mutation = list(
    RecordBeg = ~as.numeric(RecordBeg)
  ),
  id = "RecBeg_num"
)
```


## Encoding of remaining covariates

## Weighting

### By interest
As where are actually only interested in approximating $\mathbb{E}[|\text{Exposure} = 1]$, we ...
$f:[0,1]\longrightarrow\mathbb{R}_{>=0}$

While weighting is normally done to fit models on imbalanced datasets, the same method should allow us to prioritize a precise fit on the ... of interest while hopefully still extracting some information from the remaining dataset.

$$w_i = f_{k,l}(x_i):= \frac{\exp((x_i-l)\cdot k)}{1+\exp((x_i-l)\cdot k)} $$

```{r, echo=FALSE}
x<- seq(from = 0,
        to = 1,
        length.out = 101)

sigmoid <- function(x,k = 12){
  inner <- exp((x-0.5)*k)
  return(inner/(1+inner))
}

ggplot()+
  geom_line(
    aes(x = x,
        y = sigmoid(x))
  )
```
This weighting ... for $k=12$
```{r, eval = FALSE}
## Creates a weighting using exposure and sets it as the weigthing
po_add_weighting<-
  po("mutate",
     id = "create_weight",
     mutation =list(
       weights =~exp((Exposure-0.5)*12)/(1+exp((Exposure-0.5)*12))
       ))%>>%
  po("colroles",
     id = "set_weight",
     new_role = list(
       weights = "weight"
       ))
```

### By frequency

An alternative weighting approach more enlign with standard methods would be to weight to tackle the imbaling with respect to the sign of `ClaimAmount`. A quick 

```{r}
train |> 
  group_by(sign(ClaimAmount))|>
  summarize(rel_freq = n()/nrow(train))|>
  mutate(reciprocal_rel_freq = 1/rel_freq)
```
Thus the vast majority of observations have `ClaimAmount` equal to zero. bla... bla.. reciprocal frequency

# Choosing Learners and Encoding

For training an algorithm we decided to focus on three different learners, we chose to focus on a glmnet, gradient boost and random forest. We also considered implement some k-nearest neighbours algorithms, but decided against it due to the amount of features we have in the data.

## Encoding strategy

For a more fair comparison of our algorithms we decided on a common encoding strategy across all the learners. We begin with some simple baseline graphs, with no weighting and no custom encoding. Next we create some graphs with our own encoder for the `SocioCateg` variable. We then run these again with both our `interest` and our `frequency` weighting.

```{r, eval=FALSE}
Dummy_lrn_NW <- po("encode") %>>% po_RecBeg_num %>>% po_RecEnd_rc %>>% po("scale") %>>% lrn_obj |> as_learner()
Target_lrn_NW <- po("encodeimpact") %>>% po_RecBeg_num %>>% po_RecEnd_rc %>>% po("scale") %>>% lrn_obj |> as_learner()
Dummy_lrn_SocCat_NW <- po_SocCat_int %>>% po("encode") %>>% po_RecBeg_num %>>% po_RecEnd_rc %>>% po("scale") %>>% lrn_obj |> as_learner()
Target_lrn_SocCat_NW <- po_SocCat_int %>>% po("encodeimpact") %>>% po_RecBeg_num %>>% po_RecEnd_rc %>>% po("scale") %>>% lrn_obj |> as_learner()
```


## Glmnet

## Gradient Boost

## Random Forest
For the random forest implementation we used the ranger package. First when creating our random forest learners we decided that since the random forest learnes standard hyperparameters usually produce good results tuning the learners wasn't necessary. 

As for our encoding strategy here we decided as a baseline to implement algorithms using Dummy encoding, target encoding and glmm encoding. For each of these three encoders we had to different versions.

```{r, eval=FALSE}
Dummy_rforest <- po("encode") %>>% po_add_weighting %>>% po_RecBeg_num %>>% po_RecEnd_rc %>>% lrn("regr.ranger") |> as_learner()

Target_rforest <- po("encodeimpact") %>>% po_add_weighting %>>% po_RecBeg_num %>>% po_RecEnd_rc %>>% lrn("regr.ranger") |> as_learner()

glmm_rforest <- po("encodelmer", affect_columns = selector_type("factor")) %>>% po_add_weighting %>>%  po_RecBeg_num %>>% po_RecEnd_rc %>>% lrn("regr.ranger") |> as_learner() 

Custom_rforest_dummy <- po_add_weighting %>>% po_SocCat_int %>>% po_VehAge_num %>>% po_VehPrice_int %>>% po("encode") %>>% po_RecBeg_num %>>% po_RecEnd_rc %>>% lrn("regr.ranger")|> as_learner()

Custom_rforest_target <- po_add_weighting %>>% po_SocCat_int %>>% po_VehAge_num %>>% po_VehPrice_int %>>% po("encodeimpact") %>>% po_RecBeg_num %>>% po_RecEnd_rc %>>% lrn("regr.ranger")|> as_learner()

Custom_rforest_glmm <- po_add_weighting %>>% po_SocCat_int %>>% po_VehAge_num %>>% po_VehPrice_int %>>% po("encodelmer", affect_columns = selector_type("factor")) %>>% po_RecBeg_num %>>% po_RecEnd_rc %>>% lrn("regr.ranger")|> as_learner()
```

We then ran a benchmark to compare how the different factor encodings performed with the random forest. We can see the perfomance of our algorithms in the graph below.

```{r}
load("RforestBM.RData")
ggplot(rforest_BM$aggregate(list(msr("regr.mse"),
                                 msr("time_train"))))+
  geom_point(mapping = aes(x=time_train, y=regr.mse, color= learner_id))+
  geom_hline(mapping = aes(yintercept = regr.mse, color = learner_id),
             linetype = "dashed")+
  xlab("time")+
  ylab("Mean Squared Error")
```
First thing to note is that we do not have a point for the pure glmm encoding, this is because when we ran the benchmark with that learner we got an error we couldn't fix. But by looking at the plot it is just the simple target encoding which has performs the best, furthermore it's also the fastest.
